---
title: "Regression Models - Course Project"
author: "dillonchewwx"
date: "17/02/2021"
output: html_document
---

# Executive Summary

In this report, the `mtcars` dataset would be used to answer the following two questions:

1. Is an automatic or manual transmission better for MPG (miles per gallon)?
2. Quantify the MPG difference between automatic and manual transmissions.

Through a t-test, we have sufficient evidence (p=0.0014) to conclude that cars with manual transmissions provide better gas mileage than automatic transmission. Using regression models, a linear model shows that manual transmission provide 7.25 mpg more than automatic transmissions, but this single variable only accounts for 36% of the variation. By fitting additional variables to the model such as `disp`, `cyl`, `hp` and `wt`, the improvement was only 1.55 mpg, but this model accounts for 82.7% of the variation. 

# Exploratory Data Analysis
We will begin by loading the `mtcars` dataset and examining it. 
```{r}
library(datasets)
data(mtcars)
head(mtcars)
str(mtcars)
summary(mtcars$mpg)
```
It is noted that the transmission variable is stored in the column named `am` with 0 = automatic and 1 = manual - let's change the values for plotting.
```{r}
library(tidyverse)
mtcars2<-mtcars %>%
    mutate(am=ifelse(am==0, "Automatic", "Manual"))
mtcars2$am
```
Let's check the normality of the data.
```{r}
library(rstatix)
mtcars2 %>%
    group_by(am) %>%
    shapiro_test(mpg)
```
Since p > 0.05 for both Automatic and Manual transmission groups, we fail to reject the null hypothesis of the distributions being normal. We can proceed to use t-tests for the comparison of the mean MPG. 

Let's see the boxplot of `mpg` with `am`. 
```{r fig.height=5, fig.width=10}
library(ggpubr)
ggplot(mtcars2, aes(x=am, y=mpg, fill=am)) +
    geom_boxplot(outlier.shape=NA) +
    geom_jitter(height=0, width=0.1) + 
    theme_bw() +
    theme(legend.position="none", plot.title = element_text(hjust = 0.5)) +
    labs(x="Transmission", y="mpg (Miles/(US) gallon", title="Type of Transmission vs mpg") +
    stat_compare_means(method="t.test", label.x=1, label.y.npc=0.9)
```
From the boxplots and subsequent t-test (p=0.0014), we have sufficient evidence to conclude that cars with manual transmissions provide better gas mileage than automatic transmission.

# Regression Models
In this section, we will attempt to use regression models to quantify the MPG difference between automatic and manual transmissions. To start, we can try a linear model to fit the variable `am` to the outcome `mpg`.
```{r}
fit1<-lm(mpg~am, mtcars)
summary(fit1)
```
Here, we see that the mean MPG for automatic cars is 17.1, while manual transmission provides 7.25 mpg more. However, the $R^2$ value suggests that transmission only accounts for 33.8% of the total variance, and thus a multivariate model might be a better fit to the data. Nonetheless, the p-value of < 0.05 for `am` suggests that there is a linear correlation with `mpg`. 

Lets check the correlation of the other variables to mpg. 
```{r fig.height=5, fig.width=10}
library(ggcorrplot)
corr<-cor(mtcars)
p.mat<-cor_pmat(mtcars)
ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE, p.mat=p.mat, insig="blank")
```
From the correlation data, we observe that the variables drat and vs has moderately strong positive correlations with mpg (~0.6), while disp, cyl, hp and wt has strong negative correlations with mpg (~-0.8). Furthermore, we also note that all the abovementioned variables are significantly correlated with mpg. We shall move forward with multivariate models and begin by doing a nested model testing.
```{r}
fit<-lm(mpg~., data=mtcars)
fit2<-update(fit, mpg~am+drat)
fit3<-update(fit, mpg~am+drat+vs)
fit4<-update(fit, mpg~am+disp)
fit5<-update(fit, mpg~am+disp+cyl)
fit6<-update(fit, mpg~am+disp+cyl+hp)
fit7<-update(fit, mpg~am+disp+cyl+hp+wt)
fit8<-update(fit, mpg~am+drat+vs+disp+cyl+hp+wt)
fit9<-update(fit, mpg~am+cyl)
fit10<-update(fit, mpg~am+cyl+hp)
fit11<-update(fit, mpg~am+cyl+hp+wt)
fit12<-update(fit, mpg~am+hp)
fit13<-update(fit, mpg~am+hp+wt)
fit14<-update(fit, mpg~am+wt)
anova(fit, fit1, fit2, fit3, fit4, fit5, fit6, fit7, fit8, fit9, fit10, fit11, fit12, fit13, fit14)
```
From the nested model fit, we observe that the model with Df=1, lowest RSS, and significant p-value is model 8 with mpg ~ am + disp + cyl + hp + wt.
```{r}
summary(fit7)
```
From the summary, we see that the model explains 82.7% of the variation as given by the $R^2$ value, and manual transmissions result in a 1.55 mpg increase over automatic transmissions.  

# Appendix - Residual plots
```{r}
par(mfrow = c(2,2))
plot(fit7)
```