---
title: "CourseProject"
author: "dillonchewwx"
date: "17/02/2021"
output: html_document
---

# Executive Summary

In this report, the mtcars dataset would be used to answer the following two questions:

1. Is an automatic or manual transmission better for MPG (miles per gallon)?
2. Quantify the MPG difference between automatic and manual transmissions.

Using regression models, we observe ...

# Exploratory Data Analysis
We will begin by loading the mtcars dataset and examining it. 
```{r}
library(datasets)
data(mtcars)
head(mtcars)
str(mtcars)
summary(mtcars$mpg)
```
It is noted that the transmission variable is stored in the column named am with 0 = automatic and 1 = manual - let's change the values for plotting.
```{r}
library(tidyverse)
mtcars2<-mtcars %>%
    mutate(am=ifelse(am==0, "Automatic", "Manual"))
mtcars2$am
```
Let's check the normality of the data.
```{r}
library(rstatix)
mtcars2 %>%
    group_by(am) %>%
    shapiro_test(mpg)
```
Since p > 0.05 fo both Automatic and Manual transmission groups, we fail to reject the null hypothesis of the distributions being normal. We can proceed to use t-tests for the comparison of the mean mpg. 

Let's see the boxplot of mpg with transmission. 
```{r fig.height=5, fig.width=10}
library(ggpubr)
ggplot(mtcars2, aes(x=am, y=mpg, fill=am)) +
    geom_boxplot(outlier.shape=NA) +
    geom_jitter(height=0, width=0.1) + 
    theme_bw() +
    theme(legend.position="none", plot.title = element_text(hjust = 0.5)) +
    labs(x="Transmission", y="mpg (Miles/(US) gallon", title="Type of Transmission vs mpg") +
    stat_compare_means(method="t.test", label.x=1, label.y.npc=0.9)
```
From the boxplots and subsequent t-test (p=0.0014), we have sufficient evidence to conclude that cars with manual transmissions provide better gas mileage than automatic transmission.

# Regression Models
(Redo This Part with Linear Models and Multivariate Models)
In this section, we will attempt to use regression models to quantify the MPG difference between automatic and manual transmissions. Noting that transmission is binary - either automatic or manual, we shall perform a logistic regression with variables am as the outcome and mpg as the factor. The model would take the form of: 
$$Pr(AM_i|MPG_i, b_0, b_1)=\frac{exp(b_0+b_1MPG_i)}{1+exp(b_0+b_1MPG_i)}$$ 
or 
$$log(\frac{Pr(AM_i|MPG_i, b_0, b_1)}{1-Pr(AM_i|MPG_i, b_0, b_1)}=b_0+b_1MPG_i$$ 
where:

* $AM_i$ - Automatic or Manual
* $MPG_i$ - Miles per (US) Gallon
* $b_0$ - log probability of Manual if MPG is 0
* $b_1$ - increase in log probability of Manual for each additional MPG (compared to 0). 

We will now attempt to fit the data using a logistic regression model.
```{r fig.height=5, fig.width=10}
logmdl<-glm(am~mpg, "binomial", mtcars)
summary(logmdl)
exp(logmdl$coeff)
```
From the calculated odds, we note that the probability of Manual for 0 MPG is 0.001, while the increase in probability for each mpg is 1.36%.

We shall also plot the fitted values.
```{r}
ggplot(mtcars, aes(x=mpg, y=logmdl$fitted.values)) +
    geom_point() +
    labs(x="mpg (Miles/(US) gallon)", y="Probability of Manual", title="Probablity of Manual vs mpg") +
    theme_bw()
```
From the graph, 

